---
title: "Analysis Report"
output: html_notebook
---
```{r include=FALSE}
install.packages("kableExtra")
library(kableExtra)
```
This is a report generated to summarize the major results from the analysis. It is not intended to be a full picture, rather a jumping off point. This report will reference the files displayed in case you would like to examine them further.

### Cutoffs Calculated 
This table comes from cutoffs.csv
```{r}
cutoffs = read.csv("cutoffs.csv",sep = ',',strip.white = TRUE)
#cutoffs = knitr::kable(cutoffs)
cutoffs %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```

### Dada 2 Stats
This table was constructed from data in the stats-dada2.qzv object
```{r}
dadastat = read.csv("dada2_stats.tsv",sep = '\t',strip.white = TRUE)
#dadastat_t = dadastat %>% kbl() %>% kable_classic(full_width = F)
dadastat %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive")) 
```

### Sampling Depth
This table was constructed from data in sampling_depth.csv
```{r}
depth = read.csv("sampling_depth.csv", sep=",")
depth %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```

### Alpha Diversity

## Eveneness
This data comes from the file evenness-group-significance.qzv
```{r}
evenness_vals = read.csv("./evenness/metadata.tsv", sep='\t')
evenness_vals %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```

```{r}
evenness_stats = read.csv("./evenness/kruskal-wallis-pairwise-time.csv")

evenness_stats %>% kbl() %>% kable_styling(bootstrap_options = c("condensed","responsive"))
```
This diversity measurement is focused on highlighting _ in the dataset

## Faith
This data comes from the file faith-pd-group-significance.qzv.
```{r}
faith_vals = read.csv("./faith/metadata.tsv", sep='\t')
faith_vals %>% kbl() %>% kable_styling(bootstrap_options = c("striped","condensed","responsive"))
```
```{r}
faith_stats = read.csv("./faith/kruskal-wallis-pairwise-time.csv")
faith_stats %>% kbl() %>% kable_styling(bootstrap_options = c("condensed","responsive"))
```
This diversity measurement is focused on highlighting _ in the dataset

### Beta diversity core-metrics-results/unweighted_unifrac_emperor.qzv 
```{r}


```
