---
title: "Analysis Report"
output: html_notebook
---
```{r include=FALSE}
install.packages("kableExtra")
library(kableExtra)
```
This is a report generated to summarize the major results from the analysis. It is not intended to be a full picture, rather a jumping off point. This report will reference the files displayed in case you would like to examine them further.

### Cutoffs Calculated 
This table comes from cutoffs.csv
```{r}
cutoffs = read.csv("cutoffs.csv",sep = ',',strip.white = TRUE)
#cutoffs = knitr::kable(cutoffs)
cutoffs
```

### dada 2 stats
This table was constructed from data in the stats-dada2.qzv object
```{r}
dadastat = read.csv("dada2_stats.tsv",sep = '\t',strip.white = TRUE)
#dadastat_t = dadastat %>% kbl() %>% kable_classic(full_width = F)
dadastat_t
```

### sampling depth
This table was constructed from data in sampling_depth.csv
```{r}
depth = read.csv("sampling_depth.csv", sep=",")
depth
```

### Alpha diversity core-metrics-results/evenness-group-significance.qzv AND faith-pd-group-significance.qzv
```{r}

```
### Beta diversity core-metrics-results/unweighted_unifrac_emperor.qzv 
```{r}


```
